---
title: "R Notebook"
output: html_notebook
---

This notebook describes our second model, which manually fits the past data into a seasonal arima model.

```{r}
# importing libraries
library(ggplot2)
library(forecast)
library(tseries)
library(cowplot)
library(Kendall)
library(lubridate)
library(dplyr)
```


```{r}
load_daily <- read.csv(file= "../Data/Processed/daily_processed.csv", header = TRUE)
daily_load <- ts(load_daily$daily.sum, start = c(2006,1,1), frequency = 365)
```


```{r warning=FALSE}
# plotting ACF and PACF
acf_plot1 <- autoplot(Acf(daily_load, lag=1000, plot = FALSE), ylin=c(-0.5,1), main="ACF plot for daily load")
pacf_plot1 <- autoplot(Pacf(daily_load, lag=1000, plot = FALSE), ylin=c(-0.5,1), main="PACF plot for daily load")
acf_plot2 <- autoplot(Acf(daily_load, lag=40, plot = FALSE), ylin=c(-0.5,1), main="ACF plot for daily load")
pacf_plot2 <- autoplot(Pacf(daily_load, lag=40, plot = FALSE), ylin=c(-0.5,1), main="PACF plot for daily load")

plot_grid(acf_plot1, pacf_plot1, acf_plot2, pacf_plot2, nrow = 2)
```

We can observe the exponential decay of the ACF plot and the cut-off point for PACF is at lag=3. 

```{r}
# performing ADF test
print(adf.test(daily_load, alternative = "stationary"))
```

No stocastic trend is observed.

```{r}
# performing seasonal Mann Kendall test
summary(SeasonalMannKendall(daily_load))
```

Deterministic trend is observed.

```{r}
print(ndiffs(daily_load))

# differencing the series
diff_load <- diff(daily_load, lag=1, differences = 1)
```

```{r}
acf_plot_diff <- autoplot(Acf(diff_load, lag=1000, plot = FALSE), ylin=c(-0.5,1), main="ACF plot for differenced daily load")
pacf_plot_diff <- autoplot(Pacf(diff_load, lag=1000, plot = FALSE), ylin=c(-0.5,1), main="PACF plot for differenced daily load")

plot_grid(acf_plot_diff, pacf_plot_diff)
```

Signnificant ACF value is observed around lag=365, thus P=0, Q=1, D=1.
Signicicant cut-off for both ACF and PACF plots, thus p=1, q=1, d=0

```{r}
# Fitting the model into SARIMA(1,0,1)(0,1,1)
SARIMA_Model <- Arima(daily_load, order = c(1,0,1), seasonal=c(0,1,1), include.constant = TRUE)
summary(Model_SARIMA)
```


```{r}
load_daily$month <- month(load_daily$date)
load_daily$year <- year(load_daily$date)
load_Jan <- load_daily %>% filter(month == 1)
load_Feb <- load_daily %>% filter(month == 2)
load_JanFeb <- rbind(load_Jan, load_Feb)
load_JanFeb <- load_JanFeb %>%
  arrange(date) %>% 
  select(date, daily.sum) 

load_JanFeb <- load_JanFeb[-237, ]
```


```{r}
load_JanFeb_ts <- ts(load_JanFeb$daily.sum, start = c(2006,1), frequency = 59)
autoplot(load_JanFeb_ts)
acf_JanFeb <- autoplot(Acf(load_JanFeb_ts, lag = 200, plot = FALSE), ylin=c(-0.5,1), main = "ACF plot for data in Jan and Feb")
pacf_JanFeb <- autoplot(Pacf(load_JanFeb_ts, lag = 200, plot = FALSE), ylin=c(-0.5,1), main = "PACF plot for data in Jan and Feb")

plot_grid(acf_JanFeb, pacf_JanFeb)
```

```{r}
plot(decompose(load_JanFeb_ts))
```

```{r}
JanFeb_train <- load_JanFeb %>% top_n(284)
train_ts <- ts(JanFeb_train$daily.sum, start=c(2006,1), frequency = 59)
JanFeb_test <- load_JanFeb %>% top_n(284-355)
```

```{r}
SARIMA_autofit <- auto.arima(train_ts)
print(SARIMA_autofit)
```

```{r}
SARIMA_forecast <- forecast(object = SARIMA_autofit, h=71)
plot(SARIMA_forecast)
```


```{r}
score <- accuracy(SARIMA_forecast$mean, JanFeb_test$daily.sum)
print(score)
```

```{r}

autoplot(load_JanFeb_ts) +
    autolayer(SARIMA_forecast,series="SARIMA",PI=FALSE)
    ylab("Electricity Price") 
```

```{r}
SARIMA_autofit_final <- auto.arima(load_JanFeb_ts)
print(SARIMA_autofit_final)

forecasted <- forecast(object = SARIMA_autofit_final, h=59)
plot(forecasted)

predicted_load <- as.matrix(forecasted$mean)
```

```{r}
autofit_model <- data.frame(date = seq(from = as.Date("2011-01-01"), to = as.Date("2011-02-28"), by = "day"),load=round(predicted_load/24))

write.csv(autofit_model, file= "../Data/Result/autofit_model.csv", row.names = FALSE)
```

